### 2015-04-28 ApTranscriptome_TR.Rmd
Changed baseline temperature to 25C (the rearing temp) from 19.5C. No qualitative changes to results.

# get expression at optimum for A22 transcripts

Ap.high.Ac.nr.opt.exp <- Ap.high.Ac.nr$A22.opt

@@

- compare profiles of the temperature of maximum expression for thermally-responsive transcripts
- compare basal expression at optimal temperature for thermally-responsive genes
- compare degree of uregulation to constitutive expression
- compare the critical temperature of transcript upregulation, T~on~, between species for transcripts induced in response to high or low temperatures 
- compare the critical temperature of transcript downregulation, T~off~, between species for transcripts induced in response to high or low temperatures

@@

Genes upregulated in response to thermal stress in one species may have greater basal levels of expression in the other species that experiences those stressful conditions more often. To test this hypothesis, we compared expression levels near the optimal temperature (19.5C) between the two species for genes in that are either in the 'High' or 'Low' expression group in the other species. Specifically, do genes upregulated at high temperatures in *A22* have greater basal expression at optimal temperatures in *Ar*?

@@

Consistent with expectation, there is greater expression at 19.5C for *Ar* than *A22* transcripts for the set of transcripts that are transcripts that are up-regulated at high temperatures in *A22*. Note that A22 had the larger library size so if this was due to TPM not correctly accounting for differences in reads between samples, we would expect to see a positive instead of negative value here.

@@

I extract the **High** transcripts for *A. picea* and plot the log ratio of inducibility in *A. picea* over inducibility in *A. carolinensis*, against the log ratio of constitutive expression in *A. carolinensis* against expression in *A. picea*. According to the *genetic assimilation hypothesis* there should be a positive relationship between these ratios. That is, greater constitutive expression in *A. carolinensis* should have greater inducibility towards **High** temperatures in *A. picea*.

@@

# ratio of upregulation
Ap.response.type2.A22.high$A22.Ar.upreg <- Ap.response.type2.A22.high$A22.upreg / Ap.response.type2.A22.high$Ar.upreg

# ratio of constitutive expression
Ap.response.type2.A22.high$A22.Ar.constitutive <- Ap.response.type2.A22.high$A22.opt / Ap.response.type2.A22.high$Ar.opt

@@

A22.high.lm <- lm(log(Ar.A22.constitutive) ~ log(A22.Ar.upreg), data = Ap.response.type2.A22.high)

@@

In contrast to the prediction there is a negative relationship between these ratios. That is, genes with greater upregulation of expression also have greater constitutive expression. This is also consistent with the results simply comparing constitutive expression between species for the *A. picea* **High** transcripts.

@@

```{r assimilation_Ar_low}
### Repeat for A. carolinensis *Low* transcripts
Ap.response.type2.Ar.low <- Ap.response.type2[Ap.response.type2$Transcript %in% Ar.low.transcripts.df$Transcript, ]
str(Ap.response.type2.Ar.low)

@@

```{r extract_Transcripts}
A22.high.transcripts <- Ap.response.type[which(Ap.response.type$A22.type == "High"), "Transcript"]
Ar.high.transcripts <- Ap.response.type[which(Ap.response.type$Ar.type == "High"), "Transcript"]

@@

```{r T_on_high}
# transcripts expressed at *High* and *Bimodal* temperatures in A22
A22.high.TPM.dt.sub <- resp.TPM.dt.sub.pred[J(union(A22.high.transcripts, A22.bim.transcripts), "A22")]
setkey(A22.high.TPM.dt.sub, Transcript)
str(A22.high.TPM.dt.sub)

@@

As with genes turned on at high temperatures, *A. carolinensis* **Intermediate** genes are down-regulated on at higher temperatures than *A. picea* **Intermediate** genes. In contrast, towards low temperatures, the temperature of mean temperature of downregulation was higher for *A. picea* than *A. carolinensis*.

>>>>

# get expression at optimum for A22 transcripts
Ap.high.Ac.nr.opt.exp <- Ap.high.Ac.nr$A22.opt

@@

- compare profiles of the temperature of maximum expression for thermally-responsive transcripts
- compare basal expression at optimal temperature for thermally-responsive genes
- compare degree of upregulation to constitutive expression
- compare the critical temperature of transcript upregulation, T~on~, between species for transcripts induced in response to high or low temperatures 
- compare the critical temperature of transcript downregulation, T~off~, between species for transcripts induced in response to high or low temperatures

@@

Genes upregulated in response to thermal stress in one species may have greater basal levels of expression in the other species that experiences those stressful conditions more often. To test this hypothesis, we compared expression levels at the rearing temperature (25C) between the two species for genes in that are either in the 'High' or 'Low' expression group in the other species. Specifically, do genes upregulated at high temperatures in *A22* have greater basal expression at optimal temperatures in *Ar*?

@@

Consistent with expectations, there is greater expression at 25C for *Ar* than *A22* transcripts for the set of transcripts that are transcripts that are up-regulated at high temperatures in *A22*. Note that A22 had the larger library size so if this was due to TPM not correctly accounting for differences in reads between samples, we would expect to see a positive instead of negative value here.

@@

I extract the **High** transcripts for *A. picea* and examine the log ratio of inducibility against the log ratio of constitutive expression in *A. picea* over *A. carolinensis**. According to the *genetic assimilation hypothesis* there should be a negative relationship between these ratios. That is, transcripts with greater constitutive expression in *A. carolinensis* should have greater inducibility in response to **High** temperatures in *A. picea*.

@@

# ratio of upregulation in A22 vs Ar
Ap.response.type2.A22.high$A22.Ar.upreg <- Ap.response.type2.A22.high$A22.upreg / Ap.response.type2.A22.high$Ar.upreg

# ratio of constitutive expression in Ar vs A22
Ap.response.type2.A22.high$A22.Ar.constitutive <- Ap.response.type2.A22.high$A22.opt / Ap.response.type2.A22.high$Ar.opt

@@

A22.high.lm <- lm(log(A22.Ar.constitutive) ~ log(A22.Ar.upreg), data = Ap.response.type2.A22.high)

@@

In contrast to the prediction there is a positive relationship between these ratios. That is, genes with greater upregulation of expression also have greater constitutive expression.

@@

```{r assimilation_Ar_low}
### filter for A. carolinensis *Low* transcripts
Ap.response.type2.Ar.low <- Ap.response.type2[Ap.response.type2$Transcript %in% Ar.low.transcripts.df$Transcript, ]
str(Ap.response.type2.Ar.low)

@@

```{r extract_Transcripts}
Ap.response.type <- as.data.frame(Ap.response.type)
A22.high.transcripts <-  Ap.response.type[which(Ap.response.type$A22.type == "High"), "Transcript"]
Ar.high.transcripts <- Ap.response.type[which(Ap.response.type$Ar.type == "High"), "Transcript"]

@@

```{r T_on_high}
# transcripts expressed at *High* and *Bimodal* temperatures in A22
A22.high.TPM.dt.sub <- resp.TPM.dt.sub.pred[J(union(A22.high.transcripts.df$Transcript, A22.bim.transcripts), "A22")]
setkey(A22.high.TPM.dt.sub, Transcript)
str(A22.high.TPM.dt.sub)

@@

As with genes turned on at high temperatures, *A. carolinensis* **Intermediate** genes are down-regulated on at higher temperatures than *A. picea* **Intermediate** genes. Towards low temperatures, the mean temperature of downregulation was marginally higher for *A. carolinensis* than *A. picea*.
